% ======================================================================
% preface.tex
% Copyright (c) Markus Kohm, 2008-2012
%
% This file is part of the LaTeX2e KOMA-Script bundle.
%
% This work may be distributed and/or modified under the conditions of
% the LaTeX Project Public License, version 1.3c of the license.
% The latest version of this license is in
%   http://www.latex-project.org/lppl.txt
% and version 1.3c or later is part of all distributions of LaTeX
% version 2005/12/01 or later and of this work.
%
% This work has the LPPL maintenance status "author-maintained".
%
% The Current Maintainer and author of this work is Markus Kohm.
%
% This work consists of all files listed in manifest.txt.
% ----------------------------------------------------------------------
% preface.tex
% Copyright (c) Markus Kohm, 2008-2012
%
% Dieses Werk darf nach den Bedingungen der LaTeX Project Public Lizenz,
% Version 1.3c, verteilt und/oder veraendert werden.
% Die neuste Version dieser Lizenz ist
%   http://www.latex-project.org/lppl.txt
% und Version 1.3c ist Teil aller Verteilungen von LaTeX
% Version 2005/12/01 oder spaeter und dieses Werks.
%
% Dieses Werk hat den LPPL-Verwaltungs-Status "author-maintained"
% (allein durch den Autor verwaltet).
%
% Der Aktuelle Verwalter und Autor dieses Werkes ist Markus Kohm.
%
% Dieses Werk besteht aus den in manifest.txt aufgefuehrten Dateien.
% ======================================================================

\ProvidesFile{preface.tex}[2012/03/04 Preface to the KOMA-Script book]
\translator{Markus Kohm}

% Date of the translated German file: 0000/00/00

\addchap{Preface to \KOMAScript~3\protect\footnote{This chapter is still
    missing. Translators from German to English would be welcome!}}

\emph{Sorry, but the translation of this preface is still missing. Translators
  from German to English would be welcome!}

Nevertheless you should know, that this translation of the German
\KOMAScript{} manual is a work in progress. Currently the chapters of
\autoref{part:forAuthors} and \autoref{cha:scrbase}, \autoref{cha:scrlfile},
\autoref{cha:typearea-experts}, \autoref{cha:maincls-experts},
\autoref{cha:scrlttr2-experts} are up-to-date. There may still be dead-links
and broken references at these chapters, because the referenced information in
chapters of other parts could be missing.

The descriptions in \autoref{cha:scrwfile}, \autoref{cha:tocbasic},
\autoref{cha:scrhack} aren't translations from the German manual but are
primary descriptions from the implementation of these packages. They should be
up-to-date but have another structure and may be incomplete and may break
cross references from other chapters into these.

This preface will be translated last. So, the English manual will become
up-to-date, when this preface has been translated.

\endinput


%%% Local Variables: 
%%% mode: latex
%%% mode: flyspell
%%% ispell-local-dictionary: "english"
%%% coding: us-ascii
%%% TeX-master: "../guide"
%%% End: 

